ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-02 pyked.tex
More work on Python description
Work on #40

In Python, additional functionality can be imported into a script or session by the \pybox|import|
keyword. Cantera and PyKED both must be imported into the session so that we can work with the code.
In the case of Cantera, we will use many functions from the library, so we assign it an abbreviation
(\pybox|ct|) for convenience. From PyKED, we will only be using the \pybox|ChemKED| class, so this
is all that is imported:
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED
\end{pythonbox}
%
Next, we have to load the ChemKED file and retrieve the first element of the \pybox|datapoints|
list. Although this file only encodes a single experiment, the \pybox|datapoints| attribute will
always be a list (in this case, of length 1). As mentioned previously, the elements of the
\pybox|datapoints| list are instances of the \pybox|DataPoint| class, which we store in the variable
\pybox|dp|.
%
\begin{pythonbox}
ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]
\end{pythonbox}
%
The initial temperature, pressure, and mixture composition can be read from the
instance of the \pybox|DataPoint| class. \pk{} uses instances of the Pint \pybox|Quantity| class to
store values with units, while Cantera expects a floating-point value in SI
units as input. Therefore, we use the built-in capabilities of Pint to convert
the units from those specified in the \ck{} file to SI units, and we use the \pybox|magnitude|
attribute of the \pybox|Quantity| class to take only the numerical part. We also retrieve the
initial mixture mole fractions in a format Cantera will understand:
%
\begin{pythonbox}
T_initial = dp.temperature.to('K').magnitude
P_initial = dp.pressure.to('Pa').magnitude
X_initial = dp.get_cantera_mole_fraction()
\end{pythonbox}
%
With these properties defined, we have to create the objects in Cantera that represent the physical
state of the system to be studied.

# Load the mechanism and set the initial state of the mixture
gas = ct.Solution('h2-co-mechanism.cti')
gas.TPX = T_initial, P_initial, X_initial

# Create the reactor and the outside environment
reac = ct.IdealGasReactor(gas)
env = ct.Reservoir(ct.Solution('air.xml'))
\end{pythonbox}
%
In this example, \pybox|reac| represents the contents of the RCM reaction chamber and \pybox|env|
represents the contents of the environment. \verb|air.xml| is an input file included with Cantera
that represents an average composition of air. To apply the effect of the volume trace to the
\pybox|IdealGasReactor|, a \pybox|Wall| is installed between the reactor and environment and
assigned a velocity. The \pybox|Wall| allows the environment to do work on the reactor (or vice
versa) and change the reactor's thermodynamic state; we use a \pybox|Reservoir| for the environment
because in Cantera, \pybox|Reservoir|s always have a constant thermodynamic state and composition.
Using a \pybox|Reservoir| accelerates the solution compared to using two \pybox|IdealGasReactor|s,
since the composition and state of the environment are typically not necessary for the solution of
auotignition problems. Although we do not show the details here, a reference implementation of a
class that computes a wall velocity given the volume history of the reactor is available in
CanSen~\autocite{cansen}, in the \pybox|cansen.profile.VolumeProfile| class. Then, a Cantera
\pybox|ReactorNet| can be used to advance the state through autoignition, in this case to an end
time of \SI{50}{\milli\second}:
%
\begin{pythonbox}
time = dp.volume_history.time
volume = dp.volume_history.volume
ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))

>>>>

In Python, additional functionality can be imported into a script or session by the \pybox|import|
keyword. Cantera and PyKED both must be imported into the session so that we can work with the code.
In the case of Cantera, we will use many functions from the library, so we assign it an abbreviation
(\pybox|ct|) for convenience. From PyKED, we will only be using the \pybox|ChemKED| class, so this
is all that is imported:
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED
\end{pythonbox}
%
Next, we have to load the ChemKED file and retrieve the first element of the \pybox|datapoints|
list. Although this file only encodes a single experiment, the \pybox|datapoints| attribute will
always be a list (in this case, of length 1). As mentioned previously, the elements of the
\pybox|datapoints| list are instances of the \pybox|DataPoint| class, which we store in the variable
\pybox|dp|.
%
\begin{pythonbox}
ck = ChemKED('testfile_rcm.yaml')
dp = ck.datapoints[0]
\end{pythonbox}
%
The initial temperature, pressure, and mixture composition can be read from the
instance of the \pybox|DataPoint| class. \pk{} uses instances of the Pint \pybox|Quantity| class to
store values with units, while Cantera expects a floating-point value in SI
units as input. Therefore, we use the built-in capabilities of Pint to convert
the units from those specified in the \ck{} file to SI units, and we use the \pybox|magnitude|
attribute of the \pybox|Quantity| class to take only the numerical part. We also retrieve the
initial mixture mole fractions in a format Cantera will understand:
%
\begin{pythonbox}
T_initial = dp.temperature.to('K').magnitude
P_initial = dp.pressure.to('Pa').magnitude
X_initial = dp.get_cantera_mole_fraction()
\end{pythonbox}
%
With these properties defined, we have to create the objects in Cantera that represent the physical
state of the system to be studied. In Cantera, the \pybox|Solution| class stores the thermodynamic,
kinetic, and transport data from an input file in the CTI format. After the \pybox|Solution| object
is created, we can set the initial temperature, pressure, and mole fractions using the \pybox|TPX|
attribute of the \pybox|Solution| class:

\begin{pythonbox}
# Load the mechanism and set the initial state of the mixture
gas = ct.Solution('h2-co-mechanism.cti')
gas.TPX = T_initial, P_initial, X_initial
\end{pythonbox}

With the thermodynamic and kinetic data loaded and the initial conditions defined, we  need to
install the \pybox|Solution| instance into an \pybox|IdealGasReactor| which implements the equations
for mass, energy, and species conservation.

# Create the reactor and the outside environment
reac = ct.IdealGasReactor(gas)
env = ct.Reservoir(ct.Solution('air.xml'))
\end{pythonbox}
%
In this example, \pybox|reac| represents the contents of the RCM reaction chamber and \pybox|env|
represents the contents of the environment. \verb|air.xml| is an input file included with Cantera
that represents an average composition of air. To apply the effect of the volume trace to the
\pybox|IdealGasReactor|, a \pybox|Wall| is installed between the reactor and environment and
assigned a velocity. The \pybox|Wall| allows the environment to do work on the reactor (or vice
versa) and change the reactor's thermodynamic state; we use a \pybox|Reservoir| for the environment
because in Cantera, \pybox|Reservoir|s always have a constant thermodynamic state and composition.
Using a \pybox|Reservoir| accelerates the solution compared to using two \pybox|IdealGasReactor|s,
since the composition and state of the environment are typically not necessary for the solution of
autoignition problems. Although we do not show the details here, a reference implementation of a
class that computes a wall velocity given the volume history of the reactor is available in
CanSen~\autocite{cansen}, in the \pybox|cansen.profile.VolumeProfile| class. Then, a Cantera
\pybox|ReactorNet| can be used to advance the state through autoignition, in this case to an end
time of \SI{50}{\milli\second}:
%
\begin{pythonbox}
time = dp.volume_history.time
volume = dp.volume_history.volume
ct.Wall(reac, env, velocity=VolumeProfile(time=time, volume=volume))
